How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-25 tse-paper/content.tex
new draft texts for lit discussion and chunk retrieval techniques

% TODO A and B should possibly be overlapping, otherwise there is an
% area where neither A nor B makes sense?
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

\subsection{Discussion}
\label{sec:lit-sur:discussion}
% TODO we are extracting failures in the following studies, which is
%mainly what
% literature does, too, right? Highlight here and mention later again

% TODO: this section is crucial.
% maybe we can highlight important findings, by categorizing them either
% as further subsubsections or making it bold in the text?
% TODO: find values for X and Y
Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

% TODO: this reads very hands-off and researchy.
% I don't understand the first sentence and would simply cut it.
We saw that various researchers process build logs, often because the
logs provide more or the necessary
information to the researchers~\cite{ren2018automated}.
\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large
number of builds.
Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
% The chunk retrieval techniques we investigate within this article are
% automated, apart from their configuration through training examples.

% TODO: draw a conclusion together with figure 5: if reuse were higher,
%lots of effort could be saved?
We observed very little reuse of build log analysis tools.
This can stem from the
high specialization of the developed tools in regards to supported
% TODO: build environments is used a bunch (also in the following), but
% never defined.
build environments and targeted information.

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that many studies and tools target
only a few distinct build environments (roughly, area A).
% TODO: passive ...
The majority of build environments was however only targeted by one
article each (B).
From this, we follow: (a) there is a niche for custom parsers or even
specialized plugins integrated into the build process to make
post-factum
parsing of logs redundant.
Plugins  similar to, for example, the Maven plugin to report test
executions Bell et al.
%TODO cite
wrote, still have the disadvantage of not being able to work on historic
builds.
(b) the large tail of seldomly-investigated environments would benefit
from the existence of a more generic
solution.
% TODO this is disconnected, at least here
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}

\begin{table*}[htb]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}llXll@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

% TODO CARO
% start by defining in more detail what a chunk retrieval technique is
% or: characteristics that distinguishe them / to specify them further
% (To Moritz: we have short definitions before, the first time we use
% 'chunk', this is far to detailed to be earlier than here)
% Following that we present each of the techniques in detail.

This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
In addition, we present a comparison baseline of random line
retrieval (RLR), illustrate the techniques on a concrete example
of log chunks and sketch out their implementation.

% TODO CARO this is not enough for Table 3.
%Where does it come from?
% How did it originate? Also: never single sentence abstracts
\Cref{tab:ctr} shows a comparison of the implemented techniques.
% TODO CARO move to end of characteristics section and say that it is
% summary of the techniques along the aspects introducted
% can lead reading / understanding of the future sections

% TODO QUESTION Somehow, we have to find a better place for this.
%We have been
% using the term 'chunk retrieval' beforehand already -- this
% definition now comes too late.
%The first paragraph should already
% apply to the literature mapping, the second of course only applies
% here.

\subsection{The case for chunk retrieval on build logs}

% TODO: the following paragraphs have nothing to do with a discussion

In summary, the results of the literature study show the need for
build log
analysis techniques that are automated, build-environment agnostic
% TODO who is the user?
and can be configured by a user with little effort.
% We achieve this by developing three techniques for which the
% information to be retrieved is specified through examples or keywords.

%of the literature survey results.
% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The user provides these examples and keywords, which can simply be
exchanged whenever new cases are introduced.

We propose the notion of chunk retrieval techniques which retrieve
a substring, that describes a specific information, from a build log.
Generally, these can also be used to classify build logs and therefore
cover the great majority of the type of log analysis we saw within
our survey.

We implement three chunk retrieval techniques based on approaches
described by several articles:
PBE mimics regular expressions and parsers while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad-hoc approach of
searching for specific strings or keywords within KWS\@.

We evaluate the three chunk retrieval techniques on the
\emph{LogChunks} data set because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of the techniques.
The targeted chunks are describing error messages or
``the reason the build failed,'' which was also the information
most targeted by the articles in our survey.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs.
%, i.e., techniques that do not aggregate, combine, or deducted
%information.
We call such pieces of information from build logs
\emph{chunks} and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
% The
% granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

@@

% Moritz: The following is very obvious and could apply
% to any
% % technique.
% % kinda.
% It does give a descriptive error message on no match (and
% % that is not possible in other techniques)
% TODO Caro: Can you given an example of such an error message?
% % otherwise yes, nothing interesting happens in the application of pbe
% % though for consistency we might need this section? or merge
% config/application?
% Why does it return an error message? Shouldn't it also return
% <empty string>? Does it not rather return an error message if no
% program can be synthesized?
% TODO Moritz: a regular expression can match an empty string
% and that is different from not finding a match
% it returns error messages in both cases
% TODO: Ok, but what is the point of matching on an empty string?
%(if I understand what you mean)? you would not normally define such
%an extraction?
% If this distinction is important it needs to be made more clear

% TODO CARO: make more clear that not matching is an error cause
% pattern didn't match?
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE
% Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
% Moritz: I think it is good, but the VSM is still unexplained.
%We need a
% (at least) 1-sentence introduction to it, at its first reference
% TODO Moritz: I removed all other references, this is now the first one.
% I hope I got your comment right and now the first and only reference
% is explained :)
% TODO Slightly better, but I still have a problem with this -- what
% exactly is the vector
% space model here?
% At a minimum, the text should mention what the vector is and how the
% model is used
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in this case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

>>>>

% when we argue about this later we leave out the middle so A and B
% match.
whe then say that the long tail benefits from the generic
% solution.
Yes all benefit, but I think it's simpler like this.
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

\subsection{Discussion}
\label{sec:lit-sur:discussion}

Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

% TODO we are extracting failures in the following studies, which is
%mainly what
% literature does, too, right? Highlight here and mention later again

% TODO: this section is crucial.
% maybe we can highlight important findings, by categorizing them either
% as further subsubsections or making it bold in the text?

We saw that various researchers analyze build logs, often because the
logs are the only available source for the
information they need for their studies~\cite{ren2018automated,
seo2014programmers,beller2017oops,zampetti2017open,rausch2017empirical}.

\textbf{most analysis aims to find out why the build failed}

\textbf{most techniques retrieve chunks from build logs}

\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large number of builds.
Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
\textbf{Developing automated analysis tools requires a lot of effort.}
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.

We observed \textbf{very little reuse of build log analysis tools.}
This can stem from the
high specialization of the developed tools in regards to supported
% TODO: build environments is used a bunch (also in the following), but
% never defined.
build environments and targeted information.

% TODO: draw a conclusion together with figure 5: if reuse were higher,
%lots of effort could be saved?
Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build env -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build environments are the target of
many studies and tools (roughly, area A).
The majority of build environments was however only targeted by one
article each (roughly, area B).
From this we follow, that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}

% TODO we argue this in the intro!
% problem: we have no data on this
\textbf{techniques vary and seem rather randomly applied}

\textbf{modifying the build output is not helpful}, as well as returning
precise data through other documents than the log.
This requires access to the build configuration, is also specific to
the used build environment and prohibits the analysis of
historical data.

\textbf{TODO embed custom plugin discussion}
% TODO very disconnected here
From this, we follow: 1) there is a niche for custom parsers or even
specialized plugins integrated into the build process to make
post-factum
parsing of logs redundant.
Plugins  similar to, for example, the Maven plugin to report test
executions Bell et al.
%TODO cite
wrote, have the disadvantage of not being able to work on historic
builds.
2) the large tail of seldomly-investigated environments would benefit
from the existence of a more generic
solution.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO here we sell our techniques as superior.
match in intro?
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
At the end we illustrate how the techniques operate on a concrete
example and sketch out their implementation.

\begin{table*}[htb]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}llXll@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:techniques}
\end{table*}

\subsection{The case for chunk retrieval on build logs}
% FIXME Caro
In summary, the results of the literature study show the need for
build log
analysis techniques that are \emph{automated},
\emph{build-environment agnostic}
and can be \emph{configured with little effort}.

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
The user provides these examples and keywords, which can simply be
exchanged whenever new cases are introduced.

To analyze build logs,
we propose to use chunk retrieval techniques.
These retrieve a substring,
that describes a specific information, from a build log.
Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\ by checking
which substring was retrieved.
Therefore, chunk retrieval techniques
cover the great majority of the types of log analysis we saw within
our survey.

We implement three chunk retrieval techniques based on approaches
previously described by several articles:
PBE mimics regular expressions and parsers while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad-hoc approach of
searching for specific strings or keywords within KWS\@.

To evaluate and compare the three chunk retrieval techniques
we create a data set with logs from failed builds and corresponding
chunks that describe error messages or
``the reason the build failed.''
This was also the information
most targeted by the articles in our survey
In addition to that, our data set encompasses a broad range of
build environments which enables us to measure the generic
applicability of the techniques.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
Next, we will present our definition of chunk retrieval techniques
which automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.
% and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
% The
% granularity might be adjustable by configuration.

\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.
\Cref{tab:techniques} summarizes and distinguishes the techniques
we present along the
characteristics described above.

@@

\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program.
% I hope this is fine, you can also cut this sentence if it still
% bothers you :)
If the program finds no match because the analyzed build logs
does not contain the structure defined by the training examples,
PBE returns an error message.

@@

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE.
The lines of the output strings of the training examples define the
search query.
We represent each line as a term vector where each entry counts
how often a word appears within the line.
These vectors exist in a space ith as many dimensions as different
words appear in the log text.
This Vector Space Model~\cite{schutze2008introduction} defines
text lines as more similar if they contain the same words in the same
frequencies, meaning their term vectors point in a similar direction
in the vector space.
We use the cosine function to calculate the angle between two
term vectors.
The smaller the angle between term vectors, the more similar are the
corresponding lines.
This is called \emph{cosine similarity}~\cite{korenius2007principal}.

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.
